---
title: "SDR 2023 Landscape"
author: "Connor Flynn"
date: "1/30/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(naniar)
library(plotly)
library(ggpubr)
library(leaflet)
library(rnaturalearth)
library(sf)
library(missRanger)
library(cluster)
library(factoextra)
library(randomForest)
library(rnaturalearthdata)

```


```{r}
sdr_data <- read.csv(here("data/SDR-2023-Data.csv"))
```

```{r}
sdr_data <- sdr_data %>% 
  clean_names()
```


```{r}

unique_colors <- c("green" = "darkseagreen", "orange" = "darkorange1", "red" = "coral2", "yellow" = "darkgoldenrod2")

goal_7_faceted_bar_plot <- ggplot(data = sdr_data, aes(x = goal_7_score, 
                            y = reorder(country, goal_7_score),
                            fill = goal_7_dash)) +
  geom_col(stat = "identity") +
  facet_wrap(~regions_used_for_the_sdr, scales = "free_y") +
  scale_fill_manual(values = unique_colors) +  # Specify manual fill scale
  theme_minimal() +
  theme(axis.text.y = element_text(size = 4)) +
  labs(x = "SDG 7 Score",
       y = "")

ggplotly(goal_7_faceted_bar_plot)

```

```{r}
ggplot(sdr_data, aes(x = goal_7_score, 
                     y = goal_1_score)) +
  theme_minimal() +
  geom_point() +
  geom_smooth() +
  stat_cor()
```

```{r}
goal_7_and_9_scatter_plot <- ggplot(sdr_data, aes(x = goal_7_score, 
                     y = goal_1_score,
                     color = regions_used_for_the_sdr,
                     label = country)) +
  theme_minimal() +
  geom_point() +
  scale_color_brewer(palette = "Set3")

ggplotly(goal_7_and_9_scatter_plot)

```

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
```

```{r}
world <- world %>% 
  select(name_long, iso_a3, geometry)
```


```{r}
# Rename a column in a data frame or matrix
colnames(sdr_data)[which(colnames(sdr_data) == "country_code_iso3")] <- "iso_a3"
```


```{r}
joined_df <- left_join(sdr_data, world, by = "iso_a3")
```


```{r}
world_df_joined <- st_as_sf(joined_df)
```

```{r}
world_df_joined <- st_transform(world_df_joined, "+proj=longlat +datum=WGS84")
```


```{r}
mytext <- paste(
    "Country: ", world_df_joined$country,"<br/>", 
    "Goal 7 Score: ", round(world_df_joined$goal_7_score, 2), 
    sep="") %>%
  lapply(htmltools::HTML)

leaflet(world_df_joined) %>% 
  addTiles()  %>% 
  setView( lat=10, lng=0 , zoom=2) %>%
  addPolygons(stroke = FALSE, fillOpacity = 0.5, smoothFactor = 0.5, color = ~colorQuantile("YlOrRd", goal_7_score)(goal_7_score), label = mytext)
```

Correlation matrix

```{r}
sdr_scores <- sdr_data %>%
  select(
    goal_1_score, goal_2_score, goal_3_score, goal_4_score, goal_5_score,
    goal_6_score, goal_7_score, goal_8_score, goal_9_score, goal_10_score,
    goal_11_score, goal_12_score, goal_13_score, goal_14_score, goal_15_score,
    goal_16_score, goal_17_score
  )
```



```{r}
sdr_scores_matrix <- as.matrix(sdr_scores)
```


```{r}
cor <- cor(sdr_scores_matrix, use = "complete.obs")
```

```{r}
ggcorrplot::ggcorrplot(cor, method = "circle", type = "lower", lab = TRUE)
```

goal 7 tile

```{r}
goal_seven_data <- sdr_data %>% 
  select(country, regions_used_for_the_sdr, normalized_score_sdg7_elecac, normalized_score_sdg7_cleanfuel, normalized_score_sdg7_co2twh, normalized_score_sdg7_renewcon, goal_7_dash, goal_7_trend)
```


```{r}
# Melt the data for easier plotting
# Reshape the data using pivot_longer
melted_data <- pivot_longer(goal_seven_data, cols = starts_with("normalized_score_sdg7"), 
                            names_to = "variable", values_to = "value")

# Plotting heatmap using geom_tile
ggplot(melted_data, aes(x = variable, y = country, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c() +
  facet_wrap(~regions_used_for_the_sdr, scales = "free_y") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 5),
        axis.text.y = element_text(size = 5)) +
  labs(x = "", y = "", fill = "Score") 
```
*I want dash and trend also on the x axis*


```{r}
ggplot(sdr_data, aes(x = goal_7_score, fill = regions_used_for_the_sdr)) +
  geom_histogram(color = "black") +
  theme_minimal() +
  scale_fill_viridis_d()
```




Missing


```{r}
gg_miss_var(sdr_scores)
```

```{r}
ggplot(sdr_data, aes(x = goal_7_score, 
                     y = goal_1_score,
                     color = regions_used_for_the_sdr,
                     label = country)) +
  theme_minimal() +
  geom_miss_point() +
  scale_color_brewer(palette = "Set3")


```





Cleaning and Imputation for ML
- Clustering and Random Forest Regression/Classification

```{r}
sdr_data_normalized_scores <- sdr_data %>% 
  select(country, contains("normalized_score"))
```

```{r}
gg_miss_var(sdr_data_normalized_scores, show_pct = TRUE)
```
```{r}
sdr_data_normalized_scores_less_na <- sdr_data_normalized_scores %>%
  select(where(~ sum(is.na(.))/length(.) <= 0.3))
```


```{r}
sdr_data_imputed <- missRanger(sdr_data_normalized_scores_less_na)
```

Cluster 

```{r}
sdr_data_imputed <- sdr_data_imputed %>% 
  remove_rownames %>% 
  column_to_rownames(var="country")
```


```{r}
fviz_nbclust(sdr_data_imputed, kmeans, method = "silhouette") 
```


```{r}
k2 <- kmeans(sdr_data_imputed, centers = 2)
```

```{r}
fviz_cluster(k2, data = sdr_data_imputed) +
  theme_minimal() 
```

Find key drivers of clustering (HI Appleseed Analysis)



Random Forest


```{r}
rf_matmort <- randomForest(normalized_score_sdg3_matmort ~ .,
                             data = sdr_data_imputed,
                             importance = TRUE)
```

```{r}
rf_matmort
```

```{r}
importance_df <- as.data.frame(rf_matmort$importance)
```


```{r}
importance_df_top_10 <- importance_df %>%
  rownames_to_column(var = "variable") %>% 
  slice_max(n = 10, order_by = `%IncMSE`)
```

```{r}
ggplot(importance_df_top_10, aes(x = `%IncMSE`, y = reorder(variable, `%IncMSE`))) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  theme_minimal()
  
```
Partial Dependence Plots

Week 2 Day 1 - EDA - getting to know your data, troubleshooting
Week 2 Day 2 - Bar chart, basic viz, Maps, Histogram, Bubble Plots

Week 3 Day 1 - Scatterplot w/line, correlation matrices
Week 3 Day 2 - Machine Learning, imputing data, clustering, random forest


